Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management

### 2012-09-18 chi2012.tex

Second, we conducted a quantitative, structural analysis of various popular personal data sources available today to identify potential barriers to unifying data.  Drawing upon definitions from the database integration and automatic schema matching literature, we characterised the types of heterogeneity exhibited across data sources in six domains: contacts, events, music, shopping, social networking and weather, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.

@@

%% TODO THIS IS PLACEHOLDER DATA >> 
\begin{table*}
\small
\begin{tabular}{| p{2.0cm} | p{1.5cm} | p{3.6cm} | p{2cm} | p{2cm} | l | l | l | l |}
\hline
source category  & data type & sources & API type & avg fields/rec & avg depth & term. & struct & semantic \\
\hline
social networking  & person's profile & Facebook, LinkedIn, Google+ & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
event calendars & events & Eventbrite, Google Calendar  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
retail (shopping)  & product profile & Amazon, Shopping.com, Etsy, Groupon, Zappos  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
music catalogues & songs and albums & Echonest, rdio, Soundcloud, Spotify  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
weather & forecasts & W.Underground, Weatherbug, Yahoo! Weather & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
\hline
\end{tabular}
\caption{\emph{PS.2 Results} - Analysis of heterogeneity present across data feeds, by category of feed}\label{tbl:prestudy2}

\end{table*}
%% <<

\subsection{PS.2 - Technical Challenges of Data Integration}
Our investigation of integration challenges started with identifying a set of candidate data sources for analysis.  For this, we consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} of 50 most popular data feeds measured by the number of developer API keys issued for each.  In order to get a variety of different kinds of sources represented, we first categorised feeds into 5 topic categories: social networks (people), retail (shopping), events (calendar events, happenings, etc), music, and weather. We then populated each category with 2-5 feeds each depending on feed availability, for a total of 15 total feeds.

For each feed, 3-5 data records were obtained by querying the API directly for records using popular key terms for each category's primary \emph{corresponding type}, listed in the second colum of Table \ref{tbl:prestudy2}. While many of the APIs also provided access to other kinds of data, such as customer reviews, only the primary type were obtained for each category; e.g., product records for the \emph{shopping} category.  Once data records were obtained, two complexity metrics were computed for each, followed by pairwise heterogeneity metrics.  Since, due to optional fields,  individual records could vary in the number of fields.  An alternative approach to using instance documents was to directly compare schema definitions directly; however, this approach was abandoned after an initial investigation revealed two problems with schema analysis.  First, many of the sites' schema documentation was not actually publicly available; second, that many of the schemas were not actually representative of typical data returned by the service. For example, the schema documentation for \emph{shopping.com} listed full product attribute properties, but none of the records actually returned for the products we queried contained such attributes.

>>>>

% Second, we conducted a quantitative, structural analysis of various popular personal data sources available today to identify potential barriers to effective unification of data.  Drawing upon definitions from the database integration and automatic schema matching literature, we characterised the types of heterogeneity exhibited across data sources in six domains: address books (contact info), event calendars (events), music (songs/tracks), shopping (product info) social networking (profiles) and weather (forecasts), finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.

Second, we conducted a structural analysis of various popular personal data sources available today to identify potential barriers to unifying data.  Drawing upon definitions from the database integration and automatic schema matching literature, we characterised the types of heterogeneity exhibited across data sources in six domains: contacts, events, music, shopping, social networking and weather, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.

@@

%% TODO THIS IS PLACEHOLDER DATA >> 
\begin{table*}
\small
\begin{tabular}{| p{2.0cm} | p{1.5cm} | p{3.6cm} | p{2cm} | p{2cm} | l | l | l | l |}
\hline
source category  & data type & sources & API type & avg fields/rec & avg depth & term. & struct & semantic \\
\hline
social networking  & person's profile & Facebook, LinkedIn, Google+ & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
event calendars & events & Eventbrite, Google Calendar  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
retail (shopping)  & product profile & Amazon, Shopping.com, Etsy, Groupon, Zappos  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
music catalogues & songs and albums & Echonest, rdio, Soundcloud, Spotify  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
weather & forecasts & W.Underground, Weatherbug, Yahoo! Weather & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
\hline
\end{tabular}
\caption{\emph{PS.2 Results} - Analysis of heterogeneity present across data feeds, by category of feed}\label{tbl:prestudy2}
\end{table*}
%% <<

\subsection{PS.2 - Technical Challenges of Data Integration}
Our investigation of integration challenges of available structured data feeds started with identifying a set of candidate sources to examine.  For this, we consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} of 50 most popular data feeds measured by the number of developer API keys issued for each.  To increase the variety of sources, we first categorised sources into 5 topic categories: social network services, retailers, event sources (e.g., online calendars and event sites) music, and weather, and selected 2-5 sources from each 2-5 feeds each for a total of 20 total feeds for analysis.

For each feed, 3-5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, , this was user profiles, this was done either by finding record IDs by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query otherwise.  listed in the second colum of Table \ref{tbl:prestudy2}. While many of the APIs also provided access to other kinds of data, such as customer reviews, only the primary type were obtained for each category; e.g., product records for the \emph{shopping} category.  Once data records were obtained, two complexity metrics were computed for each, followed by pairwise heterogeneity metrics described below.  An alternative approach to using instance documents was to directly compare schema definitions directly; however, many of the sites' schemas were not actually publicly available; second, that many of the schemas were not actually representative of typical data returned by the service. Thus we decided to use an instance document random sampling technique instead.
